---
title: "ESM 232 - Assignment 2"
author: "Margaux Sleckman"
date: "May 15, 2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```



###First, connect Rmd SQL chunk to SQLite database

```{r library, results='hide', message=FALSE}

library(DBI)
library(tidyverse)
library(dplyr)
if(!require(readr)) install.packages("readr")
if(!require(skimr)) install.packages("skimr")
library(stringr)
library(dbplyr)


library("RSQLite")

```


###Create database and lead data into it

```{r db_creation_load_data}

## create SQl database named parcels_db 
parcels_db <- dbConnect(RSQLite::SQLite(), dbname = "parcels.db")
parcels_db

## Load data into R as df
### Use Codes - same df as assignment 1
UseCodes <- read_delim("UseCodes.csv", delim = "|", na = "No Description Available",
                     col_types = cols_only(UseCode = "n", CodeDesc = "c"))
    ## Note: one cell is NA unknown. Will get a warning

### parcels - this dataframe was written before the analysis assignment 1 (column selection and coltype definition are est. in assignment 1) 

parcels <- read_delim("parcels.csv", delim = "|")

## Load data into parcels_db
dbWriteTable(parcels_db, "parcels", parcels, row.names = T, overwrite = T)
dbWriteTable(parcels_db, "UseCodes", UseCodes, row.names = T, overwrite = T)

##verify that the two dfs are in parcels_db
db_list_tables(parcels_db)

```


###1. What are the 10 most-frequently-occuring land uses (in desc. order)

*Results in table below*
```{sql connection=parcels_db}

SELECT parcels.usecode, UseCodes.CodeDesc, count(parcels.UseCode) AS "Count"
FROM parcels 
JOIN UseCodes ON parcels.UseCode = UseCodes.UseCode
Group By parcels.UseCode
ORDER BY Count DESC
Limit 10;


```


###2. How many acres are in agricultural preserves?
  To do this, we summed the acres for only agricultural preserve (AgPres) type land 
  We assume agri preserve is everything non-NA in agri preserve varaible
  
*single results in table below*

```{sql connection=parcels_db}

Select sum(parcels.Acreage) As "Total acreage" 
From parcels
where parcels.AgPres IS NOT NULL;


```


###3. What is the mean net assessed value per acre of the entire county?
* Assuming the entire county is a single parcel, we take the sum net_AV and divide it by the sume of a acres in the county. 
* We assume negative Net_AV values are assumed to be valid (even the -1) 

*single results in table below*

```{sql connection = parcels_db}

Select (sum(parcels.Net_AV))/(sum(parcels.Acreage)) AS "Mean Net Assessed Value"
FROM parcels;


```


###4. What is the total net assessed value of all non-taxable parcels

*single result in table below*
```{sql connection=parcels_db}

Select sum(parcels.Net_AV) As "Total Net_AV for Non-taxable" 
From parcels
where parcels.NonTaxCode IS NOT NULL;


```


###5. Who are the 10 largest property holders, by acreage?

We assume use mailing address of tax statements as a proxy for the parcel owner

*results in table*

```{sql connection = parcels_db}


SELECT parcels.Mail_Address1, parcels.Mail_Address2,  sum(parcels.Acreage) AS "TotalPropertyAcreage"
FROM parcels 
Where parcels.Mail_Address1 IS NOT NULL AND parcels.Mail_Address2 IS NOT NULL AND parcels.Acreage IS NOT NULL
Group By parcels.Mail_Address1, parcels.Mail_Address2
ORDER BY TotalPropertyAcreage DESC
Limit 10;

```
 
Interestingly, the largest property owners by acreage in Santa Barbara county have a mailing address in SF.


###6. Who are the 10 largest property holders, by net assessed value?

*results in table*

```{sql connection=parcels_db}

SELECT parcels.Mail_Address1, parcels.Mail_Address2, sum(parcels.Net_AV) AS "PropertyNet_AV"
FROM parcels 
Where parcels.Mail_Address1 IS NOT NULL AND parcels.Mail_Address2 IS NOT NULL AND parcels.Net_AV IS NOT NULL
Group By parcels.Mail_Address1, parcels.Mail_Address2
ORDER BY PropertyNet_AV DESC
Limit 10;

```

Interestingly, the largest SB county property holder by net assessed value has a mailing address in Downtown SB.

```{r}

dbDisconnect(parcels_db)

```
