---
title: "ESM232-Assignment1"
author: "Margaux Sleckman"
date: "May 7, 2019"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Import and Tidy

#### Libraries
```{r libs}

library(tidyverse)
library(dplyr)
if(!require(readr)) install.packages("readr")
if(!require(skimr)) install.packages("skimr")
library(stringr)


```
####1. Read the parcels data as-is
```{r read}

#---read in file---# 

# parcels <- read_csv("Assignment1/Santa_Barbara_County_parcels_2011.csv")
# View(head(parcels, 50))

# path <- "C:/github/ESM262/ESM262/"

getwd()

parcels_raw <- read_delim("Santa_Barbara_County_parcels_2011.csv", col_types = cols(.default = col_character()), delim = ',')

# parcels_raw <- read_csv("Santa_Barbara_County_parcels_2011.csv")

names(parcels_raw)

# View(head(parcels_raw, 50))

```

####2.Copy the following columns into a tibble
```{r subset_check_clean}

#---Parse Data ---#
parcel_select <- transmute(parcels_raw, 
                        APN = parse_number(APN), 
                        Home_address1 = Situs1,
                        Home_adress2 = Situs2, 
                        Acreage = parse_double(Acreage),
                        UseCode = UseCode, 
                        NonTaxCode = NonTaxCode, 
                        AgPres = AgPres, 
                        Land_Value = LandValue,
                        Net_Impr = parse_double(Net_Impr), 
                        Net_AV = parse_double(Net_AV), 
                        Mail_Address1 = M_Address1, 
                        Mail_Address2 = M_Address2)

# dim(parcel_select)
# skim(parcel_select)
# str(parcel_select)

parcels <- as_tibble(parcel_select)

## seems clean 

```
#### Convert  
```{r conversion_ifnecessary}

## skim through the data see if any significant issues  

lapply(parcels, class)
anyNA(lapply(parcels, unique))

skim(as.tibble(parcels[,11-12]))
skim(as.tibble(parcels[,6:10]))
skim(as.tibble(parcels[,1:5]))

```

####write delim parcels file
```{r write}

#write.csv of parcels df 

write_delim(parcels, "parcels.csv", delim = "|")

```


## Analysis Questions

###1. 10 most-frequently-occuring land uses
```{r 1}

## 10 most frequ.occuring land uses 
## load in data

UseCodes<-read_delim("UseCodes.csv", delim = "|", na = "No Description Available",
                     col_types = cols_only(UseCode = "c", CodeDesc = "c"))

## warning identified missing value here:
UseCodes[1430,]

#---Join----
if(!class(parcels$UseCode) == "character") as.character(parcels$UseCode)
if(!class(UseCodes$UseCode) == "character") as.character(UseCodes$UseCode)

class(parcels$UseCode)
class(UseCodes$UseCode)


# UseCodes$UseCode <- as.character(UseCodes$UseCode)
# parcels$UseCode <- as.character(parcels$UseCode)

parcels <- left_join(x = parcels, y = UseCodes, by = 'UseCode') 

lapply(parcels, class)
str(parcels)
# View(parcels)
# unique(parcels$UseCode)
```

```{r Solution1}

## Solution ##
parcels_toplanduses <- parcels %>% 
  group_by(UseCode, CodeDesc) %>% 
  tally() %>% 
  arrange(-n) %>%
  head(10)

parcels_toplanduses

## alternative solution 
parcels_toplanduses2 <- parcels %>% 
  count(UseCode, CodeDesc) %>% 
  arrange(-n) %>% 
  head(10)

colnames(parcels_toplanduses) <- c("UseCode", "CodeDesc", "Land Use Frequency")
colnames(parcels_toplanduses2) <- c("UseCode", "CodeDesc", "Land Use Frequency")

# View(parcels_toplanduses2)
# View(parcels_toplanduses)

## consider renameing the tally column n 
  
```

#### 2. Acres for the whole agri. preserves 
We summed the acres for only agri. preserve type land 
We assume agri preserve is everything non-NA in agri preserve varaible
```{r acres_agripres}

###Solution
## Here we assume that everything not defined by NA by the AgriPres. variable is agri. preserve land
parcels_acreage <- parcels %>% 
  filter(!is.na(AgPres))  

sum_acres <- sum(as.numeric(parcels_acreage$Acreage)) 
sum_acres


###Alternative: if we choose to define agri preserve land by the description
parcels_acreage <- parcels %>% 
  filter(str_detect(CodeDesc, "Ag preserve")) 

Sum_acres <- sum(as.numeric(parcels_acreage$Acreage))
Sum_acres

## note: the acreage is larger if we use code description or agripreserver variable as defining element for agri preserve land. 

```

#### 3. Mean net assessed value per acre of the entire county 
Assumption: Negative Net_av values are assumed to be valid (even the -1) 
```{r mean_net_AV}

## Net mean assess value
## Assume mean Net-AV per acre is the sum of al the Net AV / the total number of acres.  
mean_Net_AV <- sum(as.numeric(parcels$Net_AV)/sum(parcels$Acreage))
mean_Net_AV


```

#### 4. what is the total net assessed value of all non-taxable parcels 
We assume non-taxable parcels are indicated by non-NA values for NonTaxCode
```{r nontaxableparcels}

parcels_nontaxable <- parcels %>% 
  filter(!is.na(NonTaxCode)) %>% 
  filter(!is.na(Net_AV))
 
total_NetAV_nontaxable <-sum(as.numeric(parcels_nontaxable$Net_AV))
total_NetAV_nontaxable

```

#### 5. 10 largest property holders, by acreage?
  --> use mailing address of tax statements as a proxy for the parcel owner
 
```{r top_prop_holders}

parcels_top_acreage <- parcels %>% 
  filter(!is.na(Acreage)) %>% 
  filter(!is.na(Mail_Address1) | !is.na(Mail_Address2)) %>%
  mutate(full_address = paste0(Mail_Address1, " ", Mail_Address2)) %>% 
  # group_by(Mail_Address1, Mail_Address2) %>% 
  group_by(full_address) %>% 
  summarise(property_acreage = sum(as.numeric(Acreage))) %>% 
  arrange(-property_acreage) %>% 
  head(10)
  
parcels_top_acreage

## Interestingly, the largest property owners by acreage in Santa Barbara county have a mailing address in SF.
```
  
#### 6. 10 largest property holder, by net assessed value?

```{r largestpropertyholder_byNAV}

parcels_top_NetAV <- parcels %>% 
  filter(!is.na(Net_AV)) %>% 
  filter(!is.na(Mail_Address1) | !is.na(Mail_Address2)) %>% 
  mutate(full_address = paste0(Mail_Address1, " ", Mail_Address2)) %>% 
  group_by(full_address) %>% 
  summarise(property_NetAV = sum(as.numeric(Net_AV))) %>% 
  arrange(-property_NetAV) %>% 
  head(10)

parcels_top_NetAV


## The largest SB county property holder by net assessed value has a mailing address in Downtown SB.  
```

```{r question7}


## What is the air-speed velocity of an unladen swallow?

## It is an African or European Swallow!!!

```

