---
title: "ESM232-assignment3"
author: "Margaux Sleckman"
date: "May 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# what if we design new highways that have lower rolling coefficients 
# We can reduce the rolling coefficient by 50%
# Or we can reduce the mean speed to 80 km/h (still with 10% standard deviation)
# Calculate mean power for both (assuming the same car probabilities above)
# Which is better?

```{r get_function}

source("autopower.R")

```


```{r overall_setup}
#3Crolling_less_50pc reps. is 50% of 0.015, the default value for crolling in the autopower() function
crolling_less_50pc = 0.0075 

##Conversion of km/hr to m/s - conversion factor = 0.277
mean_speed = 100 * 0.277    #100km/hr to m/s
mean_speed_low = 80 * 0.277 #80km/hr to ms

## Samples form generation of random speeds following normal distribution
nsample = 10000
nsample_high = 100000

## Generation of random speeds for raondom generation
speeds = rnorm(mean = mean_speed, sd=0.1*mean_speed, n=nsample)
speeds_2 = rnorm(mean = mean_speed_low, sd=0.1*mean_speed_low, n=nsample)


## What does this look like?
summary(speeds)

```

###1. What is the mean autopower generation (W) when we have a new rolling/friction coefficient: 
```{r setup_for_1}

##Create empty DF
results_ms = data.frame(speed=speeds, power=NA)

head(results_ms)

##Generate 3 possible cars with respectvie area and mass
possible_cars_ms = data.frame(mass=c(31000,45000,38000), area = c(25,30,22))
possible_cars_ms$row = seq(from=1, to=nrow(possible_cars_ms))
possible_cars_ms$prob = c(0.4, 0.4, 0.2)

##Created a large random sample of the three cars   
whichcar = base::sample(possible_cars_ms$row,
                        size=nsample,
                        prob=possible_cars_ms$prob,
                        replace=TRUE)

head(whichcar)


## Populate results_ms DF with the different masses and areas of the three cars.
results_ms$mass = possible_cars_ms$mass[whichcar]
results_ms$area = possible_cars_ms$area[whichcar]
  ## again, there are only 3 unique options for mass and area

head(results_ms)

```

```{r run_autopower_w_new_crolling}


##Run autopower()with the lower crolling value
pb_new_crolling <- autopower(V=results_ms$speed, m = results_ms$mass, A = results_ms$area, crolling = crolling_less_50pc)

head(pb_new_crolling, 50)

##mean power generation with lower crolling (W)
mean(pb_new_crolling)

##OLD##
# pb_new_crolling <- sapply(crolling_less_50pc, 
#                autopower,
#         A=results_ms$area, 
#        V=results_ms$speed, 
#        m = results_ms$mass)

```

###2. What is the mean autopower generation (W) when the speed limit is reduced to 80 km/h 
```{r set_up_for2}

# summary(speeds_2)

## Create Empty for cars
results_ms_2 = data.frame(speed=speeds_2, power=NA)

head(results_ms_2)

results_ms_2$mass = possible_cars_ms$mass[whichcar]
results_ms_2$area = possible_cars_ms$area[whichcar]

head(results_ms_2)

# ##Conducting exactly as above (understandably redundant)
# possible_cars_ms = data.frame(mass=c(31000,45000,38000), area = c(25,30,22))
# possible_cars_ms$row = seq(from=1, to=nrow(possible_cars_ms))
# possible_cars_ms$prob = c(0.4, 0.4, 0.2)
# # possible_cars_ms$power = autopower(V=speed_base, A = possible_cars$area, m=possible_cars$area)
# 
# whichcar = base::sample(possible_cars_ms$row,
#                         size=nsample,
#                         prob=possible_cars_ms$prob,
#                         replace=TRUE)

# what is whichcar?

```

```{r run_autopower_w_new_speeds}

##Run autopower() with new speeds:
pb_new_mean_speeds <- autopower(A=results_ms_2$area, V=results_ms_2$speed, m = results_ms_2$mass)

head(pb_new_mean_speeds, 50)

##mean power generation with lower speed 
mean(pb_new_mean_speeds) # (W)

```

Option B appears to be a better option based on the above calculations because the energy output from a new rolling coefficient is higher (187,000 W) compard to the power required to move a vehicle at an 0 km/hr speed limit (~177,000 W).

In other words, a policy that reduced the speed limit to 80 km/hr will cause a greater drop in energy output compared to a policy that reduces the degree of friction onthe highway.
